Box Plots Using All Data

### 2024-03-22 A4V_BodySections_RNAseq/R code/A4V_Tables.Rmd
overdue update

library(plotly)
library(edgeR)
library(goseq)
library(org.Dm.eg.db)
library(rrvgo)
library(VennDiagram)

@@

@@

##write.csv(cpmdata, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv")
##write.csv(A4V.FDR, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv")
##write.csv(A4V.meancpm, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv")
##write.csv(A4V.FC, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv")

```{r}

cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv", row.names = 1)

meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv", row.names = 1)

FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv", row.names = 1)

FC = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv", row.names = 1)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

```{r echo = F}

## arc1 FBgn = 'FBgn0033926'
##spn FBgn = 'FBgn0010905'
FBgn = 'FBgn0035348'
name = ''

par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
gene = data.frame(cpm = as.numeric(cpm[FBgn,]), 
                  condition = substr(row.names(meta),1, nchar(row.names(meta))-1), 
                  group = paste(meta$Genotype, meta$Sex))

gene$condition = factor(gene$condition, levels = c("A3FH", "A9FH", "A40FH",
                                                   "S3FH", "S9FH", "S40FH",
                                                   "A3FT", "A9FT", "A40FT",
                                                   "S3FT", "S9FT", "S40FT",
                                                   "A3FA", "A9FA", "A40FA",
                                                   "S3FA", "S9FA", "S40FA",
                                                   "A3MH", "A9MH",
                                                   "S3MH", "S9MH",
                                                   "A3MT", "A9MT",
                                                   "S3MT", "S9MT",
                                                   "A3MA", "A9MA",
                                                   "S3MA", "S9MA"))
gene$group = factor(gene$group, levels = c(unique(gene$group)))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])
legend('topright',
       inset=c(-0.175,0),
       legend = c('A4V F', 'Silent F', 'A4V M', 'Silent F'), 
       fill = c('firebrick',"darkgreen","gold","dodgerblue"),
       cex = .65,
       bty = 'n',
       pt.cex = .5)

@@

##GO.Sim.PCA function
```{r echo = F, include = F}
GO.Sim.PCA = function(use.data){
    pca <- prcomp(t(use.data), scale.=TRUE) 
    pc1=scale(pca$x[,1])
    pc2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(PC1=pc1,
                        PC2=pc2,
                        sdev=pca$sdev,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group = reducedTerms[row.names(use.data), "parentTerm"],
                        Score = reducedTerms[row.names(use.data), "score"],
                        Term = reducedTerms[row.names(use.data), "term"],
                        TotalDE = GO.data[row.names(use.data), "numDEInCat"],
                        Total = GO.data[row.names(use.data), "numInCat"],
                        SigVal = signif(GO.data[row.names(use.data), "adjp"],3))

    pca.data$Group = factor(pca.data$Group, levels = unique(pca.data$Group))
    pca.data$Score = 20*(1+pca.data$Score/mean(pca.data$Score))

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 ##symbol = ~Group,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c("firebrick","gold","deepskyblue"),
                 marker = list(size = pca.data$Score,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Parent Group: ", pca.data$Group,
                                   "\nSample: ", pca.data$Term, 
                                   "\nTotal DE: ", pca.data$TotalDE, 
                                   "\nTotal in Cat.: ", pca.data$Total, 
                                   "\nFDR: ", pca.data$SigVal)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")),
           title = "Enriched GO term similarity")

  return(fig)
}

##GO20 Function
```{r echo = F, include =F}

GO20 = function(GO.data, GO.ontology){
GO.data = GO.data[GO.data$ontology == GO.ontology,
                  c("category", "term","adjp", "numDEInCat", "numInCat")]

data = GO.data[20:1,c("term","adjp","numDEInCat", "numInCat", "category")]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "term",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$term,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched GO Terms",
           margin = m)

  return(fig)
}

##KEGG20 Function
```{r, echo = F, include = F}

KEGG20 = function(data){

  data = data[1:20, c("category", "Name","adjp", "numDEInCat", "numInCat")]

data$term = data$Name
if(length(unique(data$term)) < length(data$term)){
  data$term[duplicated(data$term)] = paste(data$term[duplicated(data$term)], "2")
}  
data = data[20:1,]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "Name",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$Name,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched KEGG Terms",
           margin = m)

return(fig)

@@

```{r}

sigsA3FvMT = row.names(FDR)[FDR[,'A3FvMT']<.05]
sigsS3FvMT = row.names(FDR)[FDR[,'S3FvMT']<.05]
sigsA9FvMT = row.names(FDR)[FDR[,'A9FvMT']<.05]
sigsS9FvMT = row.names(FDR)[FDR[,'S9FvMT']<.05]

GR.cpm = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
GR.meta  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
GR.FDR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)

sigsGR.FxM = row.names(GR.FDR)[GR.FDR$GR.FxM <.05]
sigsWT.FxM = row.names(GR.FDR)[GR.FDR$WT.FxM <.05]

A3 = setdiff(sigsA3FvMT, sigsS3FvMT)
A9 = setdiff(sigsA9FvMT, sigsS9FvMT)
GR = setdiff(sigsGR.FxM, sigsWT.FxM)

##Sample titles as strings. Only fill in up to your number of selected categories
set1="A4V 3-day"
set2="A4V 9-day"
set3="G85R Larvae"
set4=""

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1=A3
s2=A9
s3=GR
s4=""

T3 = A3
T9 = A9

FC.subset = FC[, c("AvS3FH", "AvS9FH", "AvS40FH",
                   "AvS3MH", "AvS9MH",
                   "AvS3FT", "AvS9FT", "AvS40FT",
                   "AvS3MT", "AvS9MT",
                   "AvS3FA", "AvS9FA", "AvS40FA",
                   "AvS3MA", "AvS9MA")]
gr <- c("03FH", "09FH", "40FH",
        "03MH", "09MH",
        "03FT", "09FT", "40FT",
        "03MT", "09MT",
        "03FA", "09FA", "40FA",
        "03MA", "09MA")
colnames(FC.subset) = gr

pca <- prcomp(t(FC.subset), scale.=TRUE)

PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Tissue=substr(gr, 4, 4), 
                        Sex=substr(gr, 3, 3), 
                        Age=substr(gr, 1, 2),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 3, 3),
                                      substr(gr, 4, 4)),
                        pca.color = substr(gr, 3, 3),
                        pca.shape = substr(gr, 4, 4))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~pca.color,
                 symbol = ~pca.shape,
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

pca.data = cpmdata[,substr(colnames(cpmdata), nchar(colnames(cpmdata))-1, nchar(colnames(cpmdata))-1) == "H"]

remove = apply(pca.data, MARGIN = 1, sum) != 0

pca.data = pca.data[remove,]

pca <- prcomp(t(pca.data), scale.=TRUE) 
gr <- colnames(pca.data)
PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=substr(gr, 1, 1), 
                        Sex=substr(gr, nchar(gr)-2, nchar(gr)-2), 
                        Age=substr(gr, 2, nchar(gr)-3),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 1, 1),
                                      substr(gr, nchar(gr)-2, nchar(gr)-1)))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Genotype,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c(input$color1, input$color2, input$color3)[c(1:length(levels(pca.color)))],
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

pca.data = cpmdata[,substr(colnames(cpmdata), nchar(colnames(cpmdata))-1, nchar(colnames(cpmdata))-1) == "T"]

remove = apply(pca.data, MARGIN = 1, sum) != 0

pca.data = pca.data[remove,]

pca <- prcomp(t(pca.data), scale.=TRUE) 
gr <- colnames(pca.data)
PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=substr(gr, 1, 1), 
                        Sex=substr(gr, nchar(gr)-2, nchar(gr)-2), 
                        Age=substr(gr, 2, nchar(gr)-3),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 1, 1),
                                      substr(gr, nchar(gr)-2, nchar(gr)-1)))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Genotype,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c(input$color1, input$color2, input$color3)[c(1:length(levels(pca.color)))],
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

pca.data = A4V.cpm[,substr(colnames(A4V.cpm), nchar(colnames(A4V.cpm))-1, nchar(colnames(A4V.cpm))-1) == "A"]

remove = apply(pca.data, MARGIN = 1, sum) != 0

pca.data = pca.data[remove,]

pca <- prcomp(t(pca.data), scale.=TRUE) 
gr <- colnames(pca.data)
PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=substr(gr, 1, 1), 
                        Sex=substr(gr, nchar(gr)-2, nchar(gr)-2), 
                        Age=substr(gr, 2, nchar(gr)-3),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 1, 1),
                                      substr(gr, nchar(gr)-2, nchar(gr)-1)))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Genotype,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c(input$color1, input$color2, input$color3)[c(1:length(levels(pca.color)))],
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

>>>>

library(edgeR)
library(goseq)
library(org.Dm.eg.db)

@@

countdata=countdata[substr(row.names(countdata),1,2)=="FB",]

@@

##write.csv(cpmdata, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv")
##write.csv(countdata, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.countdata.csv")
##write.csv(A4V.FDR, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv")
##write.csv(A4V.meancpm, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv")
##write.csv(A4V.FC, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv")

@@

@@

### 2024-03-22 GeneExpressionBoxPlots.Rmd
overdue update

@@

##relish FBgn0014018
##attacin FBgn0041579
##cecropin FBgn0000276
##diptericin FBgn0004240
##drosocin FBgn0010388
##defensin FBgn0010385
##drosomycin FBgn0283461
##metchnikowin FBgn0014865
##sod1 FBgn0003462
##sod2 FBgn0010213
##catalase FBgn0000261
##gstd1 FBgn0001149

@@

FBgn = 'FBgn0004362'
name = ''

@@

>>>>

A4V.cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv",row.names = 1)

A4V.FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv")

@@

##relish FBgn0014018
##attacin FBgn0041579
##cecropin FBgn0000276
##diptericin FBgn0004240
##drosocin FBgn0010388
##defensin FBgn0010385
##drosomycin FBgn0283461
##metchnikowin FBgn0014865
##sod1 FBgn0003462
##sod2 FBgn0010213
##catalase FBgn0000261
##gstd1 FBgn0001149
## arc1 FBgn0033926

@@

FBgn = 'FBgn0033926'
name = ''

@@

```{r echo = F}

A4V.groups = as.vector(colnames(A4V.cpm))

A4V.groups = substr(A4V.groups,1, nchar(A4V.groups)-1)

par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn,]), 
                  condition = A4V.groups, 
                  group = paste0(substr(A4V.groups, 1,1), substr(A4V.groups, nchar(as.vector(A4V.groups))-1,nchar(as.vector(A4V.groups))-1)))

gene$condition = factor(gene$condition, levels = c("A3FH", "A9FH", "A40FH",
                                                   "S3FH", "S9FH", "S40FH",
                                                   "A3FT", "A9FT", "A40FT",
                                                   "S3FT", "S9FT", "S40FT",
                                                   "A3FA", "A9FA", "A40FA",
                                                   "S3FA", "S9FA", "S40FA",
                                                   "A3MH", "A9MH",
                                                   "S3MH", "S9MH",
                                                   "A3MT", "A9MT",
                                                   "S3MT", "S9MT",
                                                   "A3MA", "A9MA",
                                                   "S3MA", "S9MA"))
gene$group = factor(gene$group, levels = c(unique(gene$group)))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])
legend('topright',
       inset=c(-0.175,0),
       legend = c('A4V F', 'Silent F', 'A4V M', 'Silent F'), 
       fill = c('firebrick',"gold","darkgreen","dodgerblue"),
       cex = .65,
       bty = 'n',
       pt.cex = .5)
